A tool set for random number generation on GPUs in R

### 2021-11-24 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
paper1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2993 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

 month_gpu<-vclMatrix(month,type="integer")
 system.time(result_month <- clrng::fisher.sim(month_gpu, 1e6, streams=streams,
                    type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
 result_month$simNum
 result_month$counts
 result_month$p.value
 @
We got 409524 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}. \fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU. 
 <<timecomparemonth3Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
 month_stats <- as.vector(result_month$sim)
 @

@@

<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
#observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
week_GPU<-gpuR::vclMatrix(week,type="integer")
system.time(result_weekgpu<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
result_weekgpu$simNum
result_weekgpu$counts
result_weekgpu$p.value
@
The ``week'' table has a much smaller p-value: around 0.0001249672, which should require a larger number of simulations to get a more accurate p-value. With more than ten million simulations, we get 1205 cases and a p-value around 0.000120472 by \fct{fisher.sim}. \fct{stats::fisher.test} takes 91.6 seconds, while \fct{fisher.sim} takes about 4.282 seconds, the elapsed time is decreased to about 4.67\% of the time taken on CPU. 
<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
result_weekgpu$cpu = as.vector(result_weekgpu$sim)
@

@@

\begin{figure}[H]
\centering
<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE, dev='png'>>=
hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
abline(v = observed_m, col = "blue", lwd = 1.5)

hist(result_weekgpu$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
abline(v = observed_w, col = "blue", lwd = 1.5)
@
\caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''. The test statistic values of the observed tables are calculated by \fct{clrng::logfactSum} and are marked with a blue line on each plot. \label{fig4}}
\end{figure}

>>>>

 month_gpu<-vclMatrix(month,type="integer")
 system.time(result_month <- clrng::fisher.sim(month_gpu, 1e6, streams=streams,
                    type="double", returnStatistics=TRUE,  Nglobal = c(256,64)))
 result_month$threshold
 result_month$simNum
 result_month$counts
 result_month$p.value
 @
We got 409524 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about 0.40443, which is close enough to the p-value from \fct{stats::fisher.test}. \fct{clrng::fisher.sim} takes about 0.69 seconds, \pkg{clrng} accounts for 4.64\% of the elapsed time on CPU. 
 <<timecomparemonth3Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
 month_stats <- as.vector(result_month$sim)
 @

@@

<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
#observed_w= -clrng::logfactSum(week, c(16,16))/almost.1
week_GPU<-gpuR::vclMatrix(week,type="integer")
system.time(result_week<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
result_week$threshold
result_week$simNum
result_week$counts
result_week$p.value
@
The ``week'' table has a much smaller p-value: around 0.0001249672, which should require a larger number of simulations to get a more accurate p-value. With more than ten million simulations, we get 1205 cases and a p-value around 0.000120472 by \fct{fisher.sim}. \fct{stats::fisher.test} takes 91.6 seconds, while \fct{fisher.sim} takes about 4.282 seconds, the elapsed time is decreased to about 4.67\% of the time taken on CPU. 
<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
result_week$cpu = as.vector(result_week$sim)
@

@@

\begin{figure}[H]
\centering
<<fighistMonth, eval=TRUE, echo=FALSE, fig.show='hold', cache=TRUE, dev='png'>>=
hist(month_stats, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), breaks=40, ylab="Frequency", main="Month data")
abline(v = result_month$threshold, col = "blue", lwd = 1.5)

hist(result_week$cpu, xlab=TeX('-$\\sum(\\log(n_{ij}!))$'), ylab="Frequency", main="Week data")
abline(v = result_week$threshold, col = "blue", lwd = 1.5)
@
\caption{Approximate sampling distributions of the test statistics from the two examples ``Month'' and ``Week''. The test statistic values of the observed tables are calculated by \fct{clrng::logfactSum} and are marked with a blue line on each plot. \label{fig4}}
\end{figure}
